---
title: "Wordle Clone"
subtitle: "Adapted from Coding out loud, Ep 4"
author: "Hunter Habersaat"
output: github_document
---

```{r load-packages, message = FALSE}
#| label: packages
#| message: false
library(tidyverse)
library(ggkeyboard) # https://github.com/sharlagelfand/ggkeyboard
```

## Inspiration

[English keyboard heatmap](https://twitter.com/monachalabi/status/772102591694209024) by [Mona Chalabi](https://twitter.com/MonaChalabi)

## Task 1: Find the first letter of Wordle words

From https://github.com/tabatkins/wordle-list.

```{r}
#| label: wordle-dataset
#| message: false

# read wordle dataset
wordle <- read_csv("https://raw.githubusercontent.com/tabatkins/wordle-list/main/words", col_names = "word")

# Add first letter column
wordle <- wordle |> 
  mutate(first_letter = str_sub(word, start = 1, end = 1))

# Count the number of entries with a given first letter
wordle_counts <- wordle |>
  count(first_letter, sort = TRUE)

wordle_counts
```

## Task 2: Draw a keyboard

```{r}
#| label: ggkeyboard
#| message: false
ggkeyboard(tkl, layout = "iso")
```

## Task 3: Color keyboard keys based on frequency of first letter

### Color options

dark blue - "#2D669B"
light blue - "#359CA8"
green - "#B8C222"
yellow - "#EEDD49"
orange - "#F2BC4A"
dark orange - "#FB704A"
red - "#BE372C"

### Color scale

```{r}
#| label: word-counts
#| message: false

# Summarize min, max, sd, and quantiles
wordle_counts |>
  summarize(
    min = min(n),
    mean = mean(n),
    sd = sd(n),
    max = max(n),
    q20 = quantile(n, 0.20),
    q40 = quantile(n, 0.40),
    q60 = quantile(n, 0.60),
    q80 = quantile(n, 0.80)
  )

# Plot histogram of first letter popularity
ggplot(wordle_counts, aes(x = n)) +
  geom_histogram(binwidth = 100)
ggplot(wordle_counts, aes(x = first_letter, y = n)) +
  geom_col() +
  geom_hline(yintercept = 80)
```
Custom Colors

- Peach         - "#FFEBE3" - 0
- Salmon        = "#F8D3CE" - 20
- Apricot       - "#F0B8A1" - 189
- Coral         - "#E89E7F" - 325
- Terracotta    - "#DC8461" - 595
- Burnt Sienna  - "#D16A44" - 736
- Rust          - "#C65027" - > 1000

```{r}
#| label: color-freq
#| message: false

# Color coding
wordle_counts <- wordle_counts |>
  mutate(
    color = case_when(
      n == 0                  ~ "#FFEBE3",
      n <=  20                ~ "#F8D3CE",
      n > 20  & n <= 189      ~ "#F0B8A1",
      n > 189 & n <= 325      ~ "#E89E7F",
      n > 325 & n <= 595      ~ "#DC8461",
      n > 595 & n <= 1000     ~ "#D16A44",
      n > 1000  ~ "#C65027"
      ),
    first_letter = str_to_upper(first_letter)
  ) |>
  arrange(first_letter)

wordle_counts
```


```{r}
#| label: key-colors
#| message: false

# Giving keys colors
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) |>
  highlight_keys(
    keys = c("Z", "X", "C", "V"), 
    fill = c("#00FF00", "#FF00FF", "#00FFFF", "#FFA500"), 
    alpha = 0.7, 
    colour = NA
    )
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) |>
  highlight_keys(
    keys = c("C", "N", "S"), 
    fill = c("#00FF00", "#FF00FF", "#00FFFF"), 
    alpha = 0.7, 
    colour = NA
    )

```
```{r}
#| label: first-letter
#| message: false

# Keys dataset, filter for first_letters
keys <- read_csv("https://raw.githubusercontent.com/sharlagelfand/ggkeyboard/061f0d9084d76d576200e81439311983e1d50d93/data-raw/sixty_percent.csv")
wordle_counts_ordered <- keys |>
  filter(key_label %in% LETTERS) |>
  select(key) |>
  left_join(wordle_counts, by = c("key" = "first_letter"))

keys
```

```{r}
#| label: keyboard1-display
#| message: false

# Assigning the colors to keys
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) |>
  highlight_keys(
    keys = wordle_counts_ordered$key, 
    fill = wordle_counts_ordered$color, 
    alpha = 0.7, 
    color = NA
    )
```
## Custom Task: Change theme and add support for 2nd and 3rd letter frequency

```{r}
#| label: packages
#| message: false

# Adding a Theme
ggkeyboard(sixty_percent, palette = keyboard_palette("serika")) |>
  highlight_keys(
    keys = wordle_counts_ordered$key, 
    fill = wordle_counts_ordered$color, 
    alpha = 0.7, 
    color = NA
    ) 
```
## 2nd Letter
```{r}
#| label: 2nd-letter
#| message: false

# Add second letter column
wordle <- wordle |> 
  mutate(second_letter = str_sub(word, start = 2, end = 2))

# Count the number of entries with a given first letter
wordle_counts2 <- wordle |>
  count(second_letter, sort = TRUE)

# Summarize min, max, sd, and quantiles
wordle_counts2 |>
  summarize(
    min = min(n),
    mean = mean(n),
    sd = sd(n),
    max = max(n),
    q20 = quantile(n, 0.20),
    q40 = quantile(n, 0.40),
    q60 = quantile(n, 0.60),
    q80 = quantile(n, 0.80)
  )

# Assign colors to frequency
wordle_counts2 <- wordle_counts2 |>
  mutate(
    color = case_when(
      n == 0                  ~ "#FFEBE3",
      n <=  20                ~ "#F8D3CE",
      n > 20  & n <= 189      ~ "#F0B8A1",
      n > 189 & n <= 325      ~ "#E89E7F",
      n > 325 & n <= 595      ~ "#DC8461",
      n > 595 & n <= 1000     ~ "#D16A44",
      n > 1000  ~ "#C65027"
      ),
    second_letter = str_to_upper(second_letter)
  ) |>
  arrange(second_letter)

# Order keys
wordle_counts2_ordered <- keys |>
  filter(key_label %in% LETTERS) |>
  select(key) |>
  left_join(wordle_counts2, by = c("key" = "second_letter"))

# Assigning the colors to keys
ggkeyboard(sixty_percent, palette = keyboard_palette("serika")) |>
  highlight_keys(
    keys = wordle_counts2_ordered$key, 
    fill = wordle_counts2_ordered$color, 
    alpha = 0.7, 
    color = NA
    )
```

## 3rd Letter

```{r}
#| label: third-letter
#| message: false

# Add third letter column
wordle <- wordle |> 
  mutate(third_letter = str_sub(word, start = 3, end = 3))

# Count the number of entries with a given first letter
wordle_counts3 <- wordle |>
  count(third_letter, sort = TRUE)

# Summarize min, max, sd, and quantiles
wordle_counts3 |>
  summarize(
    min = min(n),
    mean = mean(n),
    sd = sd(n),
    max = max(n),
    q20 = quantile(n, 0.20),
    q40 = quantile(n, 0.40),
    q60 = quantile(n, 0.60),
    q80 = quantile(n, 0.80)
  )

# Assign colors to frequency
wordle_counts3 <- wordle_counts3 |>
  mutate(
    color = case_when(
      n == 0                  ~ "#FFEBE3",
      n <=  20                ~ "#F8D3CE",
      n > 20  & n <= 189      ~ "#F0B8A1",
      n > 189 & n <= 325      ~ "#E89E7F",
      n > 325 & n <= 595      ~ "#DC8461",
      n > 595 & n <= 1000     ~ "#D16A44",
      n > 1000  ~ "#C65027"
      ),
    third_letter = str_to_upper(third_letter)
  ) |>
  arrange(third_letter)

# Order keys
wordle_counts3_ordered <- keys |>
  filter(key_label %in% LETTERS) |>
  select(key) |>
  left_join(wordle_counts3, by = c("key" = "third_letter"))

# Assigning the colors to keys
ggkeyboard(sixty_percent, palette = keyboard_palette("serika")) |>
  highlight_keys(
    keys = wordle_counts3_ordered$key, 
    fill = wordle_counts3_ordered$color, 
    alpha = 0.7, 
    color = NA
    )
```
